Supplementary Materials

### 2023-09-08 manuscript/supplementary.tex
data dryad
new updated ms from overleaf

\documentclass[11pt]{article}
\RequirePackage{fullpage}
%\RequirePackage[font=small,labelfont=bf]{caption}
\RequirePackage{amsmath,amssymb,amsthm}
\RequirePackage{graphicx}
\RequirePackage[hidelinks]{hyperref}
\RequirePackage{subcaption}
\RequirePackage{wasysym}
\RequirePackage{listings}
\RequirePackage{authblk}
\RequirePackage{bm}
\RequirePackage{bbm}
\RequirePackage{xr}
%\RequirePackage[osf]{mathpazo}
\let\temp\rmdefault
\RequirePackage{mathpazo}
\let\rmdefault\temp

@@

% for cross references
\externaldocument{manuscript}

% line numbers:
\RequirePackage{lineno}
%\modulolinenumbers[5]
\renewcommand\linenumberfont{\normalfont\tiny\sffamily\color{black}}

@@

Here, we step through a quick derivation of \textcite{Santiago1995-hx}.
Throughout, we assume random mating, hermaphroditic individuals, and a constant
population size. The change in a neutral allele's frequency in one generation
can be partitioned into the three sources of stochasticity: the random
associations with fitness backgrounds (i.e. \emph{draft}), the non-heritable
randomness in family size, and the Mendelian noise from heterozygotes
segregating. If we let $x_{0,i} \in \{0, \nicefrac{1}{2}, 1\}$ be the frequency
of neutral alleles individual $i$ in generation 0 carries, we can partition the
random neutral allele frequency of the population ($x_t$ without the individual
index) into the underlying stochastic causes,

@@

\begin{align}
  \E(S_1 S_2) &= \E(S_1^2) (1-\kappa)(1-r) = \E(S_1^2) Q_2^2 \nonumber \\
  \E(S_1 S_3) &= \E(S_1^2) (1-\kappa)^2(1-r)^2 = \E(S_1^2) Q_3^2.
\end{align}
%
The $Q_t^2$ terms represent the inflation of variance due to autocovariance in
frequency change across generations due to weak draft. Incorporating the impact
of time-invariant autocorrelation created by draft as a rescaled rate of drift
is analogous to using the Green--Kubo relation to find the transport
coefficient in molecular dynamics with velocity autocorrelation
\parencite{Green1954-kl,Kubo1957-va}.

@@

\begin{align}
  \label{supp-eq:var-xt}
  \frac{\V(x_t - p_0)}{p_0(1-p_0)} &= \sum_{i=1}^t \frac{1}{2 N} \left(V_A Q_i^2 + \frac{V_n}{4} + \frac{1}{2} \right) \left(1-\frac{1}{2 N}\right)^{i-1}.
\end{align}
%
We will see later that with the appropriate form for $Q_t^2$, this equation
holds for polygenic selection (see Supplementary Section
\ref{supp:ne-polygenic}). Additionally, we note that since the probability of
coalescence (or, equivalently, identity by descent) is proportional to the
variance in allele frequency change \parencite{Barton2000-zg}, this equation
for the variance encodes the pairwise coalescent rate of the population through
time under weak draft. Others have found that the genealogies under purifying
selection can be characterized by such a time-dependent effective population
size \parencite{Nicolaisen2013-gv}.

\subsection{The Draft-Effective Population Size}

Now, we consider how to turn this expression for the variance in neutral allele
frequency at time $t$ under weak draft into an expression for the
\emph{draft-effective} population size analogous to drift-effective population
size given in Supplementary Equation \eqref{supp-eq:var-ne}. In an ideal
Wright--Fisher population, the instantaneous variance is $\V(\Delta x_t) =
\nicefrac{p_{t-1}(1-p_{t-1})}{2N_{d,t-1}}$. Santiago and Caballero consider the
allele frequency change at some $t$ to define their effective population size
at time $t$ (which we here are calling draft-effective population size,
$N_{d,t}$),

@@

In many cases, we care about the asymptotic $N_d$,

\begin{align}
    N_d := \lim_{t \to \infty} N_{d,t} \approx \frac{4N}{4 V_A Q_\infty^2 +V_n+2}.
\end{align}
%
Now, we turn to different expressions for $Q_t^2$ and $V_A$ under different
types of polygenic selection.

\subsection{The Draft-Effective Population Size Under Polygenic Selection}
\label{supp:ne-polygenic}

The results of the previous section assume only a single locus a recombination
fraction $r$ apart from the neutral locus is determining the coalescence rate. The
model of \textcite{Santiago1998-bs} extends this to the case where fitness is
polygenic. As they show in their original paper, this system can accommodate a
variety of different equilibrium selection systems as long as expressions for
$V_A$ and $1-\kappa$ can be worked out. Other authors have extended similar models to
more elaborate bbreeding structures \parencite{Wray1990-zf,Woolliams1993-qo}.

@@

\subsection{The Draft-Effective Population Size Under Strong Background Selection}

@@

\subsection{The Draft-Effective Population Size Under Weak and Strong Negative Selection}
\label{supp:weak-strong}

With the possibility of weak selection, deleterious alleles can drift up in
frequency and fix. Consequently, the fitness distribution is no longer
stationary, as the set of haplotypes without any deleterious mutations can be
lost due to drift. This changes the shape of the fitness distribution, and thus
the variance parameter that our model relies upon
\parencite{Gessler1995-hz,OFallon2010-my,Good2013-lp,Haigh1978-gt,Higgs1995-xc}.
So far, finding an explicit equation for the variance has been difficult due to
the ``moment-closure problem" previously mentioned. In their approximation,
\textcite{Santiago2016-mu} consider a modified additive genic variance that is
below $V_\text{BGS}$ due to the loss of variation from each fixation of a
weakly deleterious mutation. This fixation process is related to the Muller's
Ratchet, the rate $R$ of fixation of deleterious mutations per generation per
region, since each fixation directly reduces the additive genic variance of the
fitness distribution. However, in general, the rate of this ratchet is unknown
since it depends on the fitness variance (and thus higher moments of the
fitness distribution).

@@

In their 2016 paper, Santiago and Caballero derive use an approach based on
Fisher's Fundamental Theorem of Natural Selection (equation 2,
\cite{Santiago2016-mu}, equation 15A \cite{Garcia-Dorado2007-jj}). A more
formal derivation follows from \textcite{Higgs1995-xc} (though they assume a
haploid asexual system, the end result is the same up to a factor of two). In
their work, they derive equations for the change per generation for the moments
and cross-moments of the fitness distribution under multiplicative selection in
a haploid model. They find, the haploid rate of the ratchet is $r = u - s v_n$
(equation 13.3, \cite{Higgs1995-xc}), where we use lowercase to distinguish
between haploid and diploid models. Here, $v_n$ is the variance in the
\emph{number} of deleterious mutations; as long as selection is not too strong,
the haploid additive variance is $v_{A} \approx s^2 v_n$. Substituting this for
$v_n$ and rearranging, the haploid fitness variance is related to the $v_A = us
- rs$, which is identical to Santiago and Caballero's
(\citeyear{Santiago2016-mu}) equation 2, noting that the waiting time between
fixations (i.e. ratchet clicks) is $T = \nicefrac{1}{r}$. Intuitively, the
fitness variance $V_\text{BGS} = Us$ is reduced by the rate at which
deleterious mutations fix and remove fitness variation from the population.

@@

\begin{align}
  \label{supp-eq:Va}
  V_{A} = Us - 2Rs. 
\end{align}
%
where $U = 2u$ and $R=r$ is the diploid rate of the ratchet, which depends on
higher moments of the fitness distribution, but only differs from $r$ through a
rescaling of the population size. We also find that Equation \eqref{eq:Va} and
Supplementary Equation \ref{supp-eq:Va} above can also be found
through Kimura's diffusion models with a flux of mutations into infinite,
discrete sites (\citeyear{Kimura1969-jw}). Note that if selected sites are
sufficiently deleterious that they cannot fix (or the population is infinite),
$R=0$ and the additive fitness variance is identical to that under strong
background selection, $V_A = Us = V_\text{BGS}$ (equation \ref{eq:va_bgs}).
We note that while we use $V_A$ here following \textcite{Santiago2016-mu}, 
this is only the \emph{additive genic} component of additive fitness variation,
as our simulations confirm in the main text.

@@

\begin{align}
  Z_\text{WS} = 1 - \frac{V_m}{V_A} = 1 - \frac{Us^2}{(U-R(N_e))s} = 1 - \frac{Us}{U-R(N_e)}.
\end{align}
%
At equilibrium, the $N_e$ that determines fixation probability is the
asymptotic $N_{d,\infty}$. Consequently, Santiago and Caballero argue that the
asymptotic reduction in draft-effective population size uses the $Q_\infty^2$ and is
given by the following system of non-linear equations,

@@

The previous section outlined the basis of quantitative genetic linked
selection models that predict the draft-effective population size under
polygenic selection and linkage. Our work here only considers the contribution
of strongly and weakly deleterious mutations to fitness variation. In this
section, we outline more details about the statistical models used to construct
the genome-wide B and B' maps that are used in model fitting.

\subsection{Parametric Forms of Negative Linked Selection Models}

@@

In practice, to fit negative selection models, we must discretize both $\mu(s,
k(g))$ and $s$ in Supplementary Equation \eqref{supp-eq:b}. We define
$\mu_{i,k(g)}$ to be the rate of mutations entering the population per basepair
per generation with selection coefficient $s_i$ for the feature class $k(g)$ of
segment $g$. This $\mu_{i, k(g)}$ term reflects the product of the mutation
rate and the distribution of deleterious fitness effects for selection
coefficient $s_i$ and feature class $k(g)$. Then, the reduction at position $x$
is,

\begin{align}
  \label{supp-eq:b-disc}
  B(x, \Psi) = \exp \left( \sum_g^S \sum_{i}^{m_s} b(\mu_{i, \;k(g)}, s_i, L_g, r_g, d(x, g))\right).
\end{align}
%
Because it is computationally-intensive to numerically solve the system of two
non-linear equations to calculate $b$ at each segment, these are pre-computed
for an $n_s$-element grid of selection coefficients and an $n_\mu$-element grid
of mutation rates--DFE products. Likewise, when we calculate $b$ under classic
BGS models, we pre-calculate segment components across these same grids, and
combine them for each $x$.

@@

Under Santiago and Caballero's model, the correct effective population size to
gauge reduction in heterozygosity or coalescent times is given by their
equation 4 \parencite{Santiago2016-mu}. This equation considers how the
inflation factor $Q_t^2$ that sets the draft-effective population size varies
over $t$. This is because statistics such as heterozygosity depend on the
cumulative autocorrelation factor up to some time $t$, $Q_t^2$ \parencite[p.
2111]{Santiago1998-bs}. This reflects the fact that effective population size
experienced by a mutation newly associated with a fitness background is larger
than the asymptotic effective population size, since not much autocorrelation
has accumulated in the stochastic perturbations caused by drift yet. From a
backwards-time perspective, this reflects the fact that pairwise diversity
(which is approximately heterozygosity when $2N\mu \ll 1$) depends on the
pairwise coalescence rate per generation, which is not constant under draft.

@@

\begin{align*}
  n_T &= \frac{n(n-1)}{2} \\
  n_S &= {c_1 \choose 2} + {c_2 \choose 2} \\
  n_D &= n_T - n_S
\end{align*}
%
which would be stored in row $\mathbf{Y}_l = [n_S, n_D]$. Note that the
per-site $\mathbf{Y}$ handles non-polymorphic sites and missing data.
Non-polymorphic sites have $n_S = {n \choose 2}$ and $n_D = 0$, and missing
data has $n_S = n_D = 0$.

@@

\subsection{Ratchet Rate Prediction}

After rescaling each segment $N_e$ by the local predicted reduction
$\hat{B}(x)$, the ratchet rates and $B'$ values are re-calculated.

The rescaled $B'$ calculation outputs a $n_\mu \times n_s \times S$
multidimensional array $\mathbf{R}$ of ratchet rates per segment. These are
rescaled by the segment lengths $\mathbf{l}$, giving the per-basepair mutation
rate array. The per-basepair ratchet rate given maximum likelihood estimates of
$\hat{\mu}$ and $\widehat{\mathbf{W}}$ is used to predicted the ratchet rate
for each segment $g$. The DFE estimate for feature $k$ is a column vector of
$\widehat{\mathbf{W}}$, i.e. $\widehat{\mathbf{w}}_k$

\begin{align}
  \lambda_D(g) = 
\end{align}

The predicted ratchet rate for segment $g$ $\lambda_d(g)$ is based on these
values.

\subsection{Numeric methods for calculating $B$}

Depending on the number of features, this can be incredibly memory intensive.

\subsection{Optimization}

Optimization in non-linear regression is notoriously difficult both in general
(Bates) and in the specific problem of estimating the effects of genome-wide
selection \parencite{Murphy2022-sj}. We experimented with a variety of global
and local optimization approaches and parameterizations, but found that

\section{Human Genomic Data}

In order to try to ensure accurate estimation of pairwise diversity, which is a
ratio estimator that is sensitivity to its denominator, we use the complete
per-basepair genotype calls (gVCF) produced by Illumina's DRAGEN pipeline
\parencite{Illumina_Inc2020-dk}. The original samples were sequenced to 30x by
the New York Genome Center \parencite{Byrska-Bishop2022-tn}. This allows for
filtering to be applied to the entire genome at once, rather than just
variants, so the denominator does not need to be estimated separately.

@@

The gVCFs were filtered using a custom Python tool, \texttt{gvcf2counts.py} (in
\path{tools/gvcf2counts.py}), which reads the gVCFs, filters them according to
the criteria below, and outputs a Numpy \path{.npy} file of reference and
alternative allele counts for each chromosome (hereafter, ``allele counts
matrices").

@@

Linked selection theory typically assumes that focal sites are not under
purifying selection themselves. Consequently, linked selection statistical
methods often only consider the fraction of the genome not \emph{a priori}
under selective constraint \parencite[e.g. Appendix Section
3.1]{Murphy2022-sj}. Our methods follow suit, and first create a track
consisting of the union (i.e. \texttt{bedtools merge}) of: exons and coding
sequences from Ensembl Version 107
\path{Homo_sapiens.GRCh38.107.chr.gff3.gz}]{Cunningham2022-vk} and PhastCons
regions \path{phastConsElements100way.bed.gz}]{Siepel2005-wh}. Then, we remove
200 basepairs of ``slop" around each of these regions (i.e. \texttt{bedtools
slop}). We show in Supplementary Figure \ref{suppfig:neutral-tracks} that
average diversity (we only consider YRI samples for this analysis) is fairly
insensitive to the slop width, but increasing slop widths around putatively
conserved segments does lead to higher diversity estimates. This is expected,
since increasing the slop width considers a smaller subset of sites, further
away from putatively conserved regions. For our analyses we choose the 200bp
slop version, which covers 63.66\% of the autosomal reference sequence and
consists of 1.26 million contiguous genomic regions.

@@

\begin{table}
    \label{supp-table:masks}
\centering
\caption{Accessible and putatively neutral sequence mask statistics, for the criteria used in 
         the main model fits.}
\begin{tabular}{|c|cccc|}
\hline
Chromosome & Accessible (\%) & Neutral (\%) & Intersection (\%) & Total Basepairs ($\times 10^6$) \\
\hline 
1 &        39.7 &     62.0 &  15.7 &     39.09 \\
2 &        45.3 &     60.1 &  18.9 &     45.77 \\
3 &        43.2 &     59.6 &  16.9 &     33.51 \\
4 &        42.7 &     65.9 &  20.2 &     38.42 \\
5 &        42.4 &     62.2 &  18.1 &     32.86 \\
6 &        43.8 &     61.4 &  18.4 &     31.43 \\
7 &        41.9 &     64.0 &  18.6 &     29.64 \\
8 &        42.3 &     66.1 &  20.2 &     29.32 \\
9 &        36.2 &     65.4 &  14.6 &     20.21 \\
10 &        42.9 &     62.2 &  18.8 &     25.15 \\
11 &        41.3 &     61.5 &  17.0 &     22.96 \\
12 &        40.6 &     60.4 &  15.9 &     21.19 \\
13 &        38.8 &     70.7 &  18.8 &     21.50 \\
14 &        36.5 &     64.6 &  13.8 &     14.77 \\
15 &        34.4 &     66.3 &  13.5 &     13.77 \\
16 &        36.0 &     66.0 &  15.5 &     14.00 \\
17 &        37.2 &     58.8 &  14.2 &     11.82 \\
18 &        40.3 &     66.1 &  19.6 &     15.75 \\
19 &        28.1 &     67.5 &  11.5 &      6.74 \\
20 &        36.1 &     62.5 &  15.4 &      9.92 \\
21 &        29.8 &     76.9 &  17.3 &      8.08 \\
22 &        25.5 &     75.5 &  12.6 &      6.40 \\
\hline
weighted average & 40.1 & 63.6 & 17.1 & - \\
\hline
\end{tabular}
\end{table}

@@

\subsection{Annotation Models}

CADD 6\%: 6.49\% of autosomal sequence.
CADD 8\%: 8.56\% of autosomal sequence.

@@

\subsection{Table of $R^2$ Values}
\label{supp:tbl-r2}
\begin{table}
\centering

@@

\subsection{Table of $R^2$ Values (Strong Selection Grid)}
\label{supp:tbl-r2-strongsel}
\begin{table}
\centering

\caption{$R^2$ and mutation rate estimates for all primary models fit with the
    strong selection grid (up to $s=10^{-1}$). Note that the repeated values in
    the last rows are correct and due to rounding; see
    \path{notebooks/main_fits.ipynb} in the GitHub repository for more
information. $R_\text{IS}^2$ is the in-sample $R^2$, and $R_\text{LOCO}^2$ is
the leave-one-chromosome-out $R^2$. }

@@

\subsection{Derivation of Approximate Coalescence $R_\text{coal}^2$}
\label{supp:r2-coal}

@@

Since this is a rough approximation,  we set $\theta = w \pi_0 B_i$ and fix
$\rho = w \gamma \pi_0 B_i$ where $\gamma$ is the ratio of recombination to
mutation rates, $\nicefrac{r}{\mu}$. In humans, $\gamma \approx 1$, but we
explore different ratios to assess sensitivity of this calculation
(Supplementary Figure \ref{suppfig:r2_upper}).

@@

Our model is essentially a generalized nonlinear model with a binomial link
function. This is the form used by \textcite{Elyashiv2016-vt} and
\textcite{Murphy2022-sj}. These models fit the observed number of pairwise
nucleotide differences in genomic windows to the expected pairwise diversity
under some evolutionary model. We only consider BGS models here, so the mean
function for position $x$ is the product of the proportion by which BGS reduces
diversity at position $x$, $B(x)$, and genome-wide neutral diversity $\pi_0$,

>>>>

\documentclass[11pt]{article}
\RequirePackage{fullpage}
%\RequirePackage[font=small,labelfont=bf]{caption}
\RequirePackage{amsmath,amssymb,amsthm}
\RequirePackage{graphicx}
\RequirePackage[hidelinks]{hyperref}
\RequirePackage{subcaption}
\RequirePackage{wasysym}
\RequirePackage{listings}
\RequirePackage{authblk}
\RequirePackage{bm}
\RequirePackage{bbm}
\RequirePackage{xr}

\newcommand*{\addFileDependency}[1]{
\typeout{(#1)}

\IfFileExists{#1}{}{\typeout{No file #1.}}
}\makeatother

\newcommand*{\myexternaldocument}[1]{%
\externaldocument{#1}%
\addFileDependency{#1.tex}%
\addFileDependency{#1.aux}%
}

\myexternaldocument{manuscript}

%\RequirePackage[osf]{mathpazo}
\let\temp\rmdefault
\RequirePackage{mathpazo}
\let\rmdefault\temp

@@

% for cross references
%\externaldocument{manuscript}

% line numbers:
\RequirePackage[left, modulo]{lineno}
\modulolinenumbers[2]
\linenumbers

% spacing
\RequirePackage{setspace}
\onehalfspacing

@@

Here, we step through a quick derivation of \textcite{Santiago1995-hx}.
Throughout, we assume random mating, hermaphroditic individuals, and a constant
population size. The change in a neutral allele's frequency in one generation
can be partitioned into the three sources of stochasticity: the random
associations with fitness backgrounds, the non-heritable
randomness in family size, and the Mendelian noise from heterozygotes
segregating. If we let $x_{0,i} \in \{0, \nicefrac{1}{2}, 1\}$ be the frequency
of neutral alleles individual $i$ in generation 0 carries, we can partition the
random neutral allele frequency of the population ($x_t$ without the individual
index) into the underlying stochastic causes,

@@

\begin{align}
  \E(S_1 S_2) &= \E(S_1^2) (1-\kappa)(1-r) = \E(S_1^2) Q_2^2 \nonumber \\
  \E(S_1 S_3) &= \E(S_1^2) (1-\kappa)^2(1-r)^2 = \E(S_1^2) Q_3^2.
\end{align}
%
The $Q_t^2$ terms represent the inflation of variance due to autocovariance in
frequency change across generations due to linked polygenic selection. Incorporating the impact
of time-invariant autocorrelation created by linked selection as a rescaled rate of drift
is analogous to using the Green--Kubo relation to find the transport
coefficient in molecular dynamics with velocity autocorrelation
\parencite{Green1954-kl,Kubo1957-va}.

@@

\begin{align}
  \label{supp-eq:var-xt}
  \frac{\V(x_t - p_0)}{p_0(1-p_0)} &= \sum_{i=1}^t \frac{1}{2 N} \left(V_A Q_i^2 + \frac{V_n}{4} + \frac{1}{2} \right) \left(1-\frac{1}{2 N}\right)^{i-1}.
\end{align}
%
We will see later that with the appropriate form for $Q_t^2$, this equation
holds for polygenic selection (see Supplementary Section
\ref{supp:ne-polygenic}). Additionally, we note that since the probability of
coalescence (or, equivalently, identity by descent) is proportional to the
variance in allele frequency change \parencite{Barton2000-zg}, this equation
for the variance encodes the pairwise coalescent rate of the population through
time under weak polygenic linked selection. Others have found that the genealogies under purifying
selection can be characterized by such a time-dependent effective population
size \parencite{Nicolaisen2013-gv}.

\subsection{The Fitness-Effective Population Size}

Now, we consider how to turn this expression for the variance in neutral allele
frequency at time $t$ under weak linked selection into an expression for the
\emph{fitness-effective} population size analogous to drift-effective population
size given in Supplementary Equation \eqref{supp-eq:var-ne}. In an ideal
Wright--Fisher population, the instantaneous variance is $\V(\Delta x_t) =
\nicefrac{p_{t-1}(1-p_{t-1})}{2N_{d,t-1}}$. Santiago and Caballero consider the
allele frequency change at some $t$ to define their effective population size
at time $t$ (which we here are calling fitness-effective population size,
$N_{d,t}$),

@@

In many cases, we care about the asymptotic $N_f$,

\begin{align}
    N_f := \lim_{t \to \infty} N_{d,t} \approx \frac{4N}{4 V_A Q_\infty^2 +V_n+2}.
\end{align}
%
Now, we turn to different expressions for $Q_t^2$ and $V_A$ under different
types of polygenic selection.

\subsection{The Fitness-Effective Population Size Under Polygenic Selection}
\label{supp:ne-polygenic}

The results of the previous section assume only a single locus a recombination
fraction $r$ apart from the neutral locus is determining the coalescence rate. The
model of \textcite{Santiago1998-bs} extends this to the case where fitness is
polygenic. As they show in their original paper, this system can accommodate a
variety of different equilibrium selection systems as long as expressions for
$V_A$ and $1-\kappa$ can be worked out. Other authors have extended similar models to
more elaborate breeding structures \parencite{Wray1990-zf,Woolliams1993-qo}.

@@

\subsection{The Fitness-Effective Population Size Under Strong Background Selection}

@@

\subsection{The Fitness-Effective Population Size Under Weak and Strong Negative Selection}
\label{supp:weak-strong}

With the possibility of weak selection, deleterious alleles can drift up in
frequency and fix. Consequently, the fitness distribution is no longer
stationary, as the set of haplotypes without any deleterious mutations can be
lost due to drift. This changes the shape of the fitness distribution, and thus
the variance parameter that our model relies upon
\parencite{Gessler1995-hz,OFallon2010-my,Good2013-lp,Haigh1978-gt,Higgs1995-xc}.
So far, finding an explicit equation for the variance has been difficult due to
the ``moment-closure problem" previously mentioned. In their approximation,
\textcite{Santiago2016-mu} consider a modified additive genic variance that is
below $V_\text{BGS}$ due to the loss of variation from each fixation of a
weakly deleterious mutation. This fixation process is related to the Muller's
Ratchet in assexual populations, the rate $R$ of fixation of deleterious mutations per generation per
region, since each fixation directly reduces the additive genic variance of the
fitness distribution. However, in general, the rate of this ratchet is unknown
since it depends on the fitness variance (and thus higher moments of the
fitness distribution).

@@

In their 2016 paper, Santiago and Caballero derive use an approach based on
Fisher's Fundamental Theorem of Natural Selection (equation 2,
\cite{Santiago2016-mu}, equation 15A \cite{Garcia-Dorado2007-jj}). A more
formal derivation follows from \textcite{Higgs1995-xc} (though they assume a
haploid asexual system, the end result is the same up to a factor of two). In
their work, they derive equations for the change per generation for the moments
and cross-moments of the fitness distribution under multiplicative selection in
a haploid model. They find the haploid rate of the ratchet is $r = u - s v_n$
(equation 13.3, \cite{Higgs1995-xc}), where we use lowercase to distinguish
between haploid and diploid models. Here, $v_n$ is the variance in the
\emph{number} of deleterious mutations; as long as selection is not too strong,
the haploid additive variance is $v_{A} \approx s^2 v_n$. Substituting this for
$v_n$ and rearranging, the haploid fitness variance is related to the $v_A = us
- rs$, which is identical to Santiago and Caballero's
(\citeyear{Santiago2016-mu}) equation 2, noting that the waiting time between
fixations (i.e. ratchet clicks) is $T = \nicefrac{1}{r}$. Intuitively, the
fitness variance $V_\text{BGS} = Us$ is reduced by the rate at which
deleterious mutations fix and remove fitness variation from the population.

@@

\begin{align}
  \label{supp-eq:Va}
  V_{A} = Us - 2Rs. 
\end{align}
%
where $U = 2u$ and $R=r$ is the diploid rate of the ratchet, which depends on
higher moments of the fitness distribution, but only differs from $r$ through a
rescaling of the population size. We also find that Equation \eqref{eq:Va} and
Supplementary Equation \ref{supp-eq:Va} above can also be found
through Kimura's diffusion models with a flux of mutations into infinite,
discrete sites (\citeyear{Kimura1969-jw}). Note that if selected sites are
sufficiently deleterious that they cannot fix (or the population is infinite),
$R=0$ and the additive fitness variance is identical to that under strong
background selection, $V_A = Us = V_\text{BGS}$ (equation \ref{supp-eq:va_bgs}).
We note that while we use $V_A$ here following \textcite{Santiago2016-mu}, 
this is only the \emph{additive genic} component of additive fitness variation,
as our simulations confirm in the main text.

@@

\begin{align}
  Z_\text{WS} = 1 - \frac{V_m}{V_A} = 1 - \frac{Us^2}{(U-R(N_e))s} = 1 - \frac{Us}{U-R(N_e)}.
\end{align}
%
At equilibrium, the $N_e$ that determines fixation probability is the
asymptotic $N_{d,\infty}$. Consequently, Santiago and Caballero argue that the
asymptotic reduction in fitness-effective population size uses the $Q_\infty^2$ and is
given by the following system of non-linear equations,

@@

The previous section outlined the basis of quantitative genetic linked
selection models that predict the fitness-effective population size under
polygenic selection and linkage. Our work here only considers the contribution
of deleterious mutations to fitness variation. In this
section, we outline more details about the statistical models used to construct
the genome-wide B and B' maps that are used in model fitting.

\subsection{Models of Linked Selection}

@@

In practice to fit negative selection models we must discretize both $\mu(s,
k(g))$ and $s$ in Supplementary Equation \eqref{supp-eq:b}. We define
$\mu_{i,k(g)}$ to be the rate of mutations entering the population per basepair
per generation with selection coefficient $s_i$ for the feature class $k(g)$ of
segment $g$. This $\mu_{i, k(g)}$ term reflects the product of the mutation
rate and the distribution of deleterious fitness effects for selection
coefficient $s_i$ and feature class $k(g)$. Then, the reduction at position $x$
is,

\begin{align}
  \label{supp-eq:b-disc}
  B(x, \Psi) = \exp \left( \sum_g^S \sum_{i}^{m_s} b(\mu_{i, \;k(g)}, s_i, L_g, r_g, d(x, g))\right).
\end{align}
%
Because it is computationally-intensive to numerically solve the system of two
non-linear equations (Equation \ref{eq:main_eqns} in the main paper) to calculate $b$ at each segment, these are pre-computed
for an $n_s$-element grid of selection coefficients and an $n_\mu$-element grid
of mutation rates--DFE products. Likewise, when we calculate $b$ under classic
BGS models, we pre-calculate segment components across these same grids, and
combine them for each $x$.

@@

Under Santiago and Caballero's model, the correct effective population size to
gauge reduction in heterozygosity or coalescent times is given by their
equation 4 \parencite{Santiago2016-mu}. This equation considers how the
inflation factor $Q_t^2$ that sets the fitness-effective population size varies
over $t$. This is because statistics such as heterozygosity depend on the
cumulative autocorrelation factor up to some time $t$, $Q_t^2$ \parencite[p.
2111]{Santiago1998-bs}. This reflects the fact that effective population size
experienced by a mutation newly associated with a fitness background is larger
than the asymptotic effective population size, since not much autocorrelation
has accumulated in the stochastic perturbations caused by drift yet. From a
backwards-time perspective, this reflects the fact that pairwise diversity
(which is approximately heterozygosity when $2N\mu \ll 1$) depends on the
pairwise coalescence rate per generation, which is not constant under linked selection.

@@

\begin{align}
  n_T &= \frac{n(n-1)}{2} \nonumber \\
  n_S &= \binom{c_1}{2} + \binom{c_2}{2} \nonumber \\
  n_D &= n_T - n_S
\end{align}
%
which would be stored in row $\mathbf{Y}_l = [n_S, n_D]$. Note that the
per-site $\mathbf{Y}$ handles non-polymorphic sites and missing data.
Non-polymorphic sites have $n_S = \binom{n}{2}$ and $n_D = 0$, and missing
data has $n_S = n_D = 0$.

@@

\section{Analysis Details}

\subsection{Human Genomic Data}

In order to try to ensure accurate estimation of pairwise diversity, which is a
ratio estimator that is sensitivity to its denominator, we use the complete
per-basepair genotype calls (gVCF) produced by Illumina's DRAGEN pipeline for the high-coverage 1000 Genome samples
\parencite{Illumina_Inc2020-dk,Byrska-Bishop2022-tn}. The original samples were sequenced to 30x by
the New York Genome Center \parencite{Byrska-Bishop2022-tn}. This allows for
filtering to be applied to the entire genome at once, rather than just
variants, so the denominator does not need to be estimated separately.

@@

The gVCFs were filtered using a custom Python tool, \texttt{gvcf2counts.py} (available in the repository associated with the manuscript at
\path{tools/gvcf2counts.py}), which reads the gVCFs, filters them according to
the criteria below, and outputs a Numpy \path{.npy} file of reference and
alternative allele counts for each chromosome (hereafter, ``allele counts
matrices").

@@

Linked selection theory typically assumes that focal sites are not under
purifying selection themselves. Consequently, linked selection statistical
methods often only consider the fraction of the genome not \emph{a priori}
under selective constraint \parencite[e.g. Appendix Section
3.1]{Murphy2022-sj}. Our methods follow suit, and first create a track
consisting of the union (i.e. \texttt{bedtools merge}) of: exons and coding
sequences from Ensembl Version 107
\path{Homo_sapiens.GRCh38.107.chr.gff3.gz}]{Cunningham2022-vk} and PhastCons
regions \path{phastConsElements100way.bed.gz}]{Siepel2005-wh}. Then, we remove
200 basepairs of buffer around each of these regions (i.e. \texttt{bedtools
slop}). We show in Supplementary Figure \ref{suppfig:neutral-tracks} that
average diversity (we only consider YRI samples for this analysis) is fairly
insensitive to the buffer width, but increasing slop widths around putatively
conserved segments does lead to higher diversity estimates. This is expected,
since increasing the slop width considers a smaller subset of sites, further
away from putatively conserved regions. For our analyses we choose the 200bp
slop version, which covers 63.66\% of the autosomal reference sequence and
consists of 1.26 million contiguous genomic regions.

@@

\begin{table}
\centering
\caption{Accessible and putatively neutral sequence mask statistics, for the criteria used in 
         the main model fits.}
\begin{tabular}{|c|cccc|}
\hline
Chromosome & Accessible (\%) & Neutral (\%) & Intersection (\%) & Total Basepairs ($\times 10^6$) \\
\hline 
1 &        39.7 &     62.0 &  15.7 &     39.09 \\
2 &        45.3 &     60.1 &  18.9 &     45.77 \\
3 &        43.2 &     59.6 &  16.9 &     33.51 \\
4 &        42.7 &     65.9 &  20.2 &     38.42 \\
5 &        42.4 &     62.2 &  18.1 &     32.86 \\
6 &        43.8 &     61.4 &  18.4 &     31.43 \\
7 &        41.9 &     64.0 &  18.6 &     29.64 \\
8 &        42.3 &     66.1 &  20.2 &     29.32 \\
9 &        36.2 &     65.4 &  14.6 &     20.21 \\
10 &        42.9 &     62.2 &  18.8 &     25.15 \\
11 &        41.3 &     61.5 &  17.0 &     22.96 \\
12 &        40.6 &     60.4 &  15.9 &     21.19 \\
13 &        38.8 &     70.7 &  18.8 &     21.50 \\
14 &        36.5 &     64.6 &  13.8 &     14.77 \\
15 &        34.4 &     66.3 &  13.5 &     13.77 \\
16 &        36.0 &     66.0 &  15.5 &     14.00 \\
17 &        37.2 &     58.8 &  14.2 &     11.82 \\
18 &        40.3 &     66.1 &  19.6 &     15.75 \\
19 &        28.1 &     67.5 &  11.5 &      6.74 \\
20 &        36.1 &     62.5 &  15.4 &      9.92 \\
21 &        29.8 &     76.9 &  17.3 &      8.08 \\
22 &        25.5 &     75.5 &  12.6 &      6.40 \\
\hline
weighted average & 40.1 & 63.6 & 17.1 & - \\
\hline
\end{tabular}
\label{supp-table:masks}
\end{table}

@@

@@

\subsection{Table of \texorpdfstring{$R^2$}{R2} Values}
\label{supp:tbl-r2}
\begin{table}
\centering

@@

\subsection{Table of \texorpdfstring{$R^2$}{R2} Values (Strong Selection Grid)}
\label{supp:tbl-r2-strongsel}
\begin{table}
\centering

\caption{$R^2$ and mutation rate estimates for all primary models fit with the
    strong selection grid (up to $s=10^{-1}$). Note that the repeated values in
    the last rows are correct and due to rounding; see
    the main analysis notebook in the GitHub repository for more information. $R_\mathrm{IS}^2$ is the in-sample $R^2$, and $R_\mathrm{LOCO}^2$ is the leave-one-chromosome-out $R^2$. }

@@

\subsection{Derivation of Approximate Coalescence \texorpdfstring{$R_\text{coal}^2$}{R2}}
\label{supp:r2-coal}

@@

Since this is a rough approximation,  we set $\theta = w \pi_0 B_i$ and fix
$\rho = w \gamma \pi_0 B_i$ where $\gamma$ is the ratio of recombination to
mutation rates, $\nicefrac{r}{\mu}$. In humans, $\gamma \approx 1$, but we
explore different ratios to assess sensitivity of this calculation
(Supplementary Figure \ref{suppfig:drift_r2}B).

@@

Our model is essentially a generalized nonlinear model with a binomial link
function. This is the form used by \textcite{Elyashiv2016-vt} and
\textcite{Murphy2022-sj}. These models fit the observed number of pairwise
nucleotide differences in genomic windows to the expected pairwise diversity
under some evolutionary model. We only consider  models of purifying selection here, so the mean
function for position $x$ is the product of the proportion by which selection reduces
diversity at position $x$, $B(x)$, and genome-wide neutral diversity $\pi_0$,
